Flow-Directed Trace Analysis for Post-Silicon Validation

### 2015-08-16 paper.tex
graph changed, added more txt

Even though this example is conceptually simple, getting the model to correctly implement the flow specifications is not straightforward.  On the other hand, results from the trace analysis greatly help the debugging process by providing information to locate problems quickly.  For example, in an early version of the model, the trace shown in Table~\ref{table-trace} is observed.  
\begin{table*}
\caption{An observed trace of messages for trace analysis.}
\label{table-trace}
\begin{center}
\begin{tabular}{llll}
1~~(CPU2, Icon, Rd, M) & 2~~(Icon, CPU1, Rd, M) & 3~~(Icon, Mem, Rd, M) & 4~~(Mem,  Icon, DVal, -) \\
5~~(CPU1, Icon, Hit, -) & 6~~(CPU1, Icon, DVal, -) & 7~~(Icon, CPU2, Wr, M) & 8~~(Icon, Mem, Wr, M) \\
9~~(CPU2, Icon, Rd, M) & 10~(Icon, CPU1, Rd, M) & 11~(Icon, Mem, Rd, M) & 12~(CPU2, Icon, Rd, P) \\
13~(Icon, CPU1, Rd, P) & 14~(CPU1, Icon, Miss, -) & 15~(Icon, Dev, Rd, P) & 16~(CPU1, Icon, DVal, -) \\
\end{tabular}
\end{center}
\end{table*}
The trace analysis finds out that messages~$1-8$ in the trace are the results of execution of an instance of flow $F_2$ as shown in Fig.~\ref{fig-cpu-read}, and the flow execution scenario is $\{(F_{2,1}, \{p_{11}, p_{12}\})\}$.  Messages~$9-11$ are analyzed as the results of executing another instance of flow $F_2$, message~$12-13$ as the results of executing an instance of flow $F_3$ as shown in Fig.~\ref{fig-peri-read}.  The flow execution scenario after the first thirteen messages in the trace is 
\[
\{(F_{2,1}, \{p_{11}, p_{12}\}), (F_{2,2}, \{p_4, p_5\}), (F_{3,1}, \{p_4, p_3\})\}.
\]  
Message~$14$ can be  the result from executing $F_{2,2}$ or $F_{3,1}$, therefore it leads to two following flow execution scenarios:
\begin{equation}
\{(F_{2,1}, \{p_{11}, p_{12}\}), (F_{2,2}, \{p_7, p_5\}), (F_{3,1}, \{p_4, p_3\})\}
\end{equation}
\begin{equation}
\{(F_{2,1}, \{p_{11}, p_{12}\}), (F_{2,2}, \{p_4, p_5\}), (F_{3,1}, \{p_6, p_3\})\}
\end{equation}
In either scenario, after mapping message~$15$ to flow $F_{3,1}$, message~$16$ cannot be mapped to any existing flow instance or to a new flow instance.  From this inconsistent message, we know that it is generated by CPU1.  In scenario~(1), CPU1 is in the state after generating the message reporting a cache miss.  In this case, the DataValid message should not generated.  In scenario~(2), CPU1 is in the state before generating the message reporting either a cache hit or miss, and again the DataValid message should not generated.   This inconsistent message helps to locate a bug in the CPU1 model where a DataValid message is generated after either a cache hit or miss message is generated.  After fixing this bug,  in a few more iterations of analysis and debugging, the trace analysis can eventually extract all initiated flow instances in the model, all in their terminal states, thus showing that the model implements the four flows correctly.

@@

The proof concept of the proposed trace analysis method is demonstrated on a transaction level model of a SoC design from GEM5 with two ARM CPU cores, and one memory connected by an interconnect as shown in Fig 1. Since the analysis method presented in this paper is communication centric, the detailed computations of these blocks are not modeled. Instead, the modeling is focused on how they participate in flows for different system level use cases.

\begin{figure} 
\centerline{
\includegraphics[width=4in]{figures/Fig4.png}}
\caption{SoC platform structure.}
\label{fig_example1.ps}
\end{figure}

@@

In this case study, two system flows are implemented in the SoC model. They include a cache coherent read request from CPU and a cache coherent write request from CPU. In the following , a brief overview for the implemented instances of these flows is given. These flow specifications capture how messages are exchanged for different use cases. In this model, a message is defined with the following format, (Src,Dest,Cmd), where Src and Dest refer to the source and destination components of messages, Cmd refers to the operations that the destination component should perform.

\begin{figure} 
\centerline{
\includegraphics[width=3in]{figures/Fig5.png}}
\caption{Flow specification ($F_1$) of a cache coherent write operation initiated from CPU1}
\label{fig_example1.ps}
\end{figure}

The LPN as shown in Fig 2 specifies a system flow where CPU1 initiates a memory write operation. In this flow, CPU1 initiates a memory write request to L1 Cache. Next, depends on whether the required data is included in cache1, the cache1 will generate three possible responses. First, the cache1 will send read exclusive request message msg2 to interconnect if data not present in cache; Or if the data is shared by CPU0, it will send upgrade message msg16 to interconnect to disable CPU0's ownership of this block of data, else if CPU1 has exclusive right of required data, cache1 will perform the write operation and sent an response message msg15 to CPU1. Afterward, interconnect will sent request to all connected component (data cache and instruction cache of each CPU and memory). Once the interconnect get the response from either CPU0 or memory, it will generate a response message to cache1 and cache1 will generate a write response message msg14 (or msg 21 ) to CPU1. The flow is symmetric for CPU2.

The LPN specification as shown in Fig 3 captures the system flow where CPU1 initiate a memory read operation. Similar to memory write, CPU1 will first initiate a memory read request message msg1 to cache1. Cache1 can generate three possible responses. If the requested data is not in cache1, cache1 will generate an storeCondReq message msg2 to interconnect asking for data. Else if the requested data is inside of cache1 but it's shared, cache1 will generate a upgrade request message msg22 to make sure it has the newest data; Else if cache1 has and data and it's not shared, it will generate a read response message msg15 to CPU1. Afterward, interconnect will send corresponding request to all its connected component and when it received the response, generate response message to cache1 and cache1 will generate the read response message msg14 or msg34 to CPU1. This specification is also symmetric for CPU2.

\begin{figure} 
\centerline{
\includegraphics[width=3in]{figures/Fih6.png}}

\caption{Flow specification ($F_2$) of a cache coherent read operation initiated from CPU1}
\label{fig_example1.ps}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Results and Discussions}

The transaction level model of the simple system implementing the two flow specifications shown in the previous section is implemented by Gem5. We build a simple platform in Gem5 with components described in Fig 1. By running two simple executables on each CPU simultaneously, we produce a result file including all the communication between each components. The executable assigned to CPU1 is read one file three times for one letter and writing one letter for three times to the same file. CPU2's executable will do the same functionality to the same file, only difference is that CPU2 will write first and then read.

Even though this example is conceptually simple, getting the model to correctly implement the flow specifications is not straightforward. On the other hand, results from the trace analysis greatly help the debugging process by providing information to locate problems quickly. For example, the following trace is observed in Fig. 4.

\begin{figure} 
\centerline{
\includegraphics[width=4in]{figures/Fig.png}}
\caption{Example of trace result}
\end{figure}
\
The trace analysis finds out that messages 1-18 maps to two CPU0 read as
\vspace{2 mm}

>>>>

Even though this example is conceptually simple, getting the model to correctly implement the flow specifications is not straightforward.  On the other hand, results from the trace analysis greatly help the debugging process by providing information to locate problems quickly.  For example, in an early version of the model, the trace shown in Table~\ref{table-trace} is observed.  
\begin{table*}
\caption{An observed trace of messages for trace analysis.}
\label{table-trace}
\begin{center}
\begin{tabular}{llll}
1~~(CPU2, Icon, Rd, M) & 2~~(Icon, CPU1, Rd, M) & 3~~(Icon, Mem, Rd, M) & 4~~(Mem,  Icon, DVal, -) \\
5~~(CPU1, Icon, Hit, -) & 6~~(CPU1, Icon, DVal, -) & 7~~(Icon, CPU2, Wr, M) & 8~~(Icon, Mem, Wr, M) \\
9~~(CPU2, Icon, Rd, M) & 10~(Icon, CPU1, Rd, M) & 11~(Icon, Mem, Rd, M) & 12~(CPU2, Icon, Rd, P) \\
13~(Icon, CPU1, Rd, P) & 14~(CPU1, Icon, Miss, -) & 15~(Icon, Dev, Rd, P) & 16~(CPU1, Icon, DVal, -) \\
\end{tabular}
\end{center}
\end{table*}
The trace analysis finds out that messages~$1-8$ in the trace are the results of execution of an instance of flow $F_2$ as shown, and the flow execution scenario is $\{(F_{2,1}, \{p_{11}, p_{12}\})\}$.  Messages~$9-11$ are analyzed as the results of executing another instance of flow $F_2$, message~$12-13$ as the results of executing an instance of flow $F_3$ as shown in Fig.~\ref{fig-peri-read}.  The flow execution scenario after the first thirteen messages in the trace is 
\[
\{(F_{2,1}, \{p_{11}, p_{12}\}), (F_{2,2}, \{p_4, p_5\}), (F_{3,1}, \{p_4, p_3\})\}.
\]  
Message~$14$ can be  the result from executing $F_{2,2}$ or $F_{3,1}$, therefore it leads to two following flow execution scenarios:
\begin{equation}
\{(F_{2,1}, \{p_{11}, p_{12}\}), (F_{2,2}, \{p_7, p_5\}), (F_{3,1}, \{p_4, p_3\})\}
\end{equation}
\begin{equation}
\{(F_{2,1}, \{p_{11}, p_{12}\}), (F_{2,2}, \{p_4, p_5\}), (F_{3,1}, \{p_6, p_3\})\}
\end{equation}
In either scenario, after mapping message~$15$ to flow $F_{3,1}$, message~$16$ cannot be mapped to any existing flow instance or to a new flow instance.  From this inconsistent message, we know that it is generated by CPU1.  In scenario~(1), CPU1 is in the state after generating the message reporting a cache miss.  In this case, the DataValid message should not generated.  In scenario~(2), CPU1 is in the state before generating the message reporting either a cache hit or miss, and again the DataValid message should not generated.   This inconsistent message helps to locate a bug in the CPU1 model where a DataValid message is generated after either a cache hit or miss message is generated.  After fixing this bug,  in a few more iterations of analysis and debugging, the trace analysis can eventually extract all initiated flow instances in the model, all in their terminal states, thus showing that the model implements the four flows correctly.

@@

\begin{figure} 
\centerline{
\includegraphics[width=3.8in]{figures/Fig4.png}}
\caption{SoC platform structure.}
\label{SOC}
\end{figure}

The proof concept of the proposed trace analysis method is demonstrated on a transaction level model of a SoC design from GEM5 with two ARM CPU cores, each with 16KB data cache and 16KB instruction cache. One simple memory with 1GB capacity is connected with cache by an interconnect bus as shown in Fig~\ref{SOC}.  Since the analysis method presented in this paper is communication centric, the detailed computations of these blocks are not modeled. Instead, the modeling is focused on how they participate in flows for different system level use cases. Originally all the component is connected by ports in GEM5, in order to trace the communication inside the system, we added an communicate monitor between each components connection. As showed in Fig~\ref{SOC}, take CPU0 and instruction cache for example, the communicate monitor between is called icache0. When CPU0 is about to send a command (read request as example), it will send to communicate monitor first, where it will store the request information in a file, and forward the request packet to instruction cache.

By combining information from all 9 of the communicate monitors, we will have all the communication traces we need. As an virtualized SOC platform, GEM5 has two types of request, timing and atomic. Timing is what will be used in our system as it simulate delays generated by 
During our experiment, we took only timing request from the raw traces to eliminate initiation traces that are useless for our research.

@@

In this case study, two system flows are implemented in the SoC model. They include a cache coherent read request from CPU and a cache coherent write request from CPU. In the following , a brief overview for the implemented instances of these flows is given. These flow specifications capture how messages are exchanged for different use cases. In this model, a message is defined with the following format, (Src,Dest,Cmd), where Src and Dest refer to the source and destination components of messages, Cmd refers to the operations that the destination component should perform. 
\begin{figure} 
\centerline{
\includegraphics[width=3.7In]{figures/Fig5.png}}

{\footnotesize
\[
\begin{array}{llll}
msg_0: (&\mbox{ CPU1},&\mbox{writeReq},&\mbox{cache1   })\\       
msg_1: (&\mbox{ cache1},&\mbox{ readExreq },&\mbox{Bus     })\\        
msg_2: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce2 })\\  
msg_3: (&\mbox{ cache2},&\mbox{readExreq},&\mbox{cpu2         })\\   
msg_4: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce2           })\\  
msg_5: (&\mbox{ cache2},&\mbox{readExreq},&\mbox{cpu2 })\\  
msg_6: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce1       })\\     
msg_7: (&\mbox{ cache1},&\mbox{readExreq},&\mbox{cpu1           })\\  
msg_8: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ Memory })\\  
msg_9: (&\mbox{ true                                          })\\  
msg_{10}: (&\mbox{ Memory},&\mbox{ readExres},&\mbox{ Bus        })\\  
msg_{11}: (&\mbox{ cache2},&\mbox{ readExres},&\mbox{ Bus })\\  
msg_{12}: (&\mbox{ Bus},&\mbox{ readExres},&\mbox{ cache1        })\\  
msg_{13}: (&\mbox{ cache1},&\mbox{ writeRes},&\mbox{ CPU1         })\\  
msg_{14}: (&\mbox{ cache1},&\mbox{ writeRes},&\mbox{ CPU1 })\\  
msg_{15}: (&\mbox{ cache1},&\mbox{UpgradeReq             })\\  
msg_{16}: (&\mbox{ Bus},&\mbox{ UpgradeReq},&\mbox{ cahce2      })\\   
msg_{17}: (&\mbox{ Bus},&\mbox{ UpgradeReq},&\mbox{ Memory })\\  
msg_{18}: (&\mbox{ cache2},&\mbox{UpgradeRes},&\mbox{ Bus     })\\  
msg_{19}: (&\mbox{ Bus},&\mbox{ UpgradeRes},&\mbox{ cache1      })\\  
msg_{20}: (&\mbox{ cache1},&\mbox{ WriteRes},&\mbox{ CPU1 })\\  
\end{array}
\]}
\caption{Flow specification ($F_1$) of a cache coherent write operation initiated from CPU1}
\label{write-flow}
\end{figure}

The LPN as shown in Fig~\ref{write-flow} specifies a system flow where CPU1 initiates a memory write operation. In this flow, CPU1 initiates a memory write request to L1 Cache. Next, depends on whether the required data is included in cache1, the cache1 will generate three possible responses. First, the cache1 will send read exclusive request message msg2 to interconnect if data not present in cache; Or if the data is shared by CPU0, it will send upgrade message msg16 to interconnect to disable CPU0's ownership of this block of data, else if CPU1 has exclusive right of required data, cache1 will perform the write operation and sent an response message msg15 to CPU1. Afterward, interconnect will sent request to all connected component (data cache and instruction cache of each CPU and memory). Once the interconnect get the response from either CPU0 or memory, it will generate a response message to cache1 and cache1 will generate a write response message msg14 (or msg 21 ) to CPU1. The flow is symmetric for CPU2.

\begin{figure} 
\centerline{
\includegraphics[width=4in]{figures/Fih6.png}}

{\footnotesize
\[
\begin{array}{llll}
t0: (&\mbox{ CPU1},&\mbox{ReadReq},&\mbox{cache1  })\\                   
t1: (&\mbox{ cache1},&\mbox{ StoreCondreq },&\mbox{Bus })\\           
t2: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce2 })\\
t3: (&\mbox{ cache2},&\mbox{StoreCondreq},&\mbox{cpu2       })\\      
t4: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce2           })\\ 
t5: (&\mbox{ cache2},&\mbox{StoreCondreq},&\mbox{cpu2 })\\
t6: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce1     })\\       
t7: (&\mbox{ cache1},&\mbox{StoreCondreq},&\mbox{cpu1           })\\ 
t8: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ Memory })\\
t9: (&\mbox{ true                                        })\\
t10: (&\mbox{ Memory},&\mbox{ ReadRes},&\mbox{ Bus            })\\    
t11: (&\mbox{ cache2},&\mbox{ ReadRes},&\mbox{ Bus })\\
t12: (&\mbox{ Bus},&\mbox{ ReadRes},&\mbox{ cache1      })\\            
t13: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1          })\\  
t14: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1 })\\
t21: (&\mbox{ cache1},&\mbox{ LoadLockedreq },&\mbox{Bus })\\     
t22: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce2     })\\
t23: (&\mbox{ cache2},&\mbox{LoadLockedreq},&\mbox{cpu2 })\\
t24: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce2     })\\ 
t25: (&\mbox{ cache2},&\mbox{LoadLockedreq},&\mbox{cpu2     })\\
t26: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce1 })\\
t27: (&\mbox{ cache1},&\mbox{LoadLockedreq},&\mbox{cpu1       })\\
t28: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ Memory     })\\
t29: (&\mbox{ true })\\
t30: (&\mbox{ Memory},&\mbox{ ReadRes},&\mbox{ Bus       })\\     
t31: (&\mbox{ cache2},&\mbox{ ReadRes},&\mbox{ Bus    })\\
t32: (&\mbox{ Bus},&\mbox{ ReadRes},&\mbox{ cache1 })\\
t33: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1 })\\
\end{array}
\]}
\caption{Flow specification ($F_2$) of a cache coherent read operation initiated from CPU1}
\label{read-flow}

\end{figure}
The LPN specification as shown in Fig~\ref{read-flow} captures the system flow where CPU1 initiate a memory read operation. Similar to memory write, CPU1 will first initiate a memory read request message msg1 to cache1. Cache1 can generate three possible responses. If the requested data is not in cache1, cache1 will generate an storeCondReq message msg2 to interconnect asking for data. Else if the requested data is inside of cache1 but it's shared, cache1 will generate a upgrade request message msg22 to make sure it has the newest data; Else if cache1 has and data and it's not shared, it will generate a read response message msg15 to CPU1. Afterward, interconnect will send corresponding request to all its connected component and when it received the response, generate response message to cache1 and cache1 will generate the read response message msg14 or msg34 to CPU1. This specification is also symmetric for CPU2.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Results and Discussions}

The transaction level model of the simple system implementing the two flow specifications shown in the previous section is implemented by Gem5. We build a simple platform in Gem5 with components described in Fig~\ref{SOC}. By running two simple executables on each CPU simultaneously, we produce a result file including all the communication between each components. The executable assigned to CPU1 is read one file three times for one letter and writing one letter for three times to the same file. CPU2's executable will do the same functionality to the same file, only difference is that CPU2 will write first and then read.

\begin{table*}
\caption{{Example of trace result}}
\label{example-trace}
\begin{center}
{\footnotesize

\begin{tabular}{llll}
1~~ (cpu0,cache0,readreq)				&2~~ (cache0,membus,LoadLockedreq)		&3~~ (membus,mem,LoadLockedreq) 		&4~~ (membus,cache0,LoadLockedreq)\\

5~~(cache0,cpu0,LoadLockedreq)		&6~~(cpu0,cache0,readreq)		&7~~(cache0,membus,StoreCondreq)		&8~~(membus,mem,StoreCondreq)\\

9~~(membus,cache0,StoreCondreq)		&10~~(cache0,cpu0,StoreCondreq)		&11~~(membus,cache1,StoreCondreq)		&12~~(cache1,cpu1,StoreCondreq)\\

13~~(membus,cache1,StoreCondreq)		&14~~(cache1,cpu1,StoreCondreq)		&15~~(membus,cache1,LoadLockedreq)	&16~~(cache1,cpu1,LoadLockedreq)\\

17~~(membus,cache1,LoadLockedreq)	&18~~(cache1,cpu1,LoadLockedreq)		&19~~(mem,membus,readres)			&20~~(cache0,cpu0,readres)\\

20~~(membus,cache0,readres)			&22~~(mem,membus,readres)			&23~~(membus,cache0,readres)				&24~~(cache0,cpu0,readres)\\
\end{tabular}}
\end{center}
\end{table*}

Even though this example is conceptually simple, getting the model to correctly implement the flow specifications is not straightforward. On the other hand, results from the trace analysis greatly help the debugging process by providing information to locate problems quickly. For example, the following trace as shown in Table~\ref{example-trace} is observed.
\
The trace analysis finds out that messages 1-18 maps to two CPU0 read as
\vspace{2 mm}
